---
title: 'Understanding Customer Churn and Prevention Strategy'
author: Package Build
date: '2021-10-30'
slug: understanding-customer-churn-and-prevention-strategy
categories:
  - R
tags:
  - Marketing
subtitle: ''
summary: ''
authors: []
lastmod: '2021-10-30T00:03:50+09:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


## Menu 

### Appetizer
+ Motivation: Segmentation is "Divide and Conquer"
+ What is RFM and why we do feature engineering 
+ Why Campaign Response important feature for CRM?
+ Power of Quantile and Pareto Principle 
+ Robust Scoring Rules 

### Main
+ Time Series Visualisation + seasonal effect
+ Data splitting 
+ RFM Scoring 
+ Visualisation for business report 

### Desert 
+ Perspective note 
+ To take action
+ Wrap up: What I have learned
+ Ready to receive feedback  


## 1.0 library load

```{r, echo = FALSE, message=FALSE, warning=FALSE}

library(readr)
library(fs)
library(tufte)
library(forecast)
library(GGally)
library(knitr)
library(kableExtra)

library(psych)
library(skimr)
library(janitor)
library(tidyverse)
library(readxl)
library(ggridges)
library(cowplot)
library(Information)

```


```{r, echo = TRUE, message=FALSE, warning=FALSE}

library(tidyverse)
library(dplyr)
library(lubridate)
library(skimr)
library(scales)
library(psych)
library(gmodels)

library(tidyquant)
library(timetk)
library(rsample) 
library(wesanderson)
```


```{r, message=FALSE, warning=FALSE}
set.seed(42)
options(scipen = 100, digits = 3)

churn_tbl <- read_csv("/Users/seunghyunsung/Desktop/R_DS/Machine_Learning/classification/randomForest/", col_types = cols())

```


```{r, message=FALSE, warning=FALSE}
resp_df_raw <- read_csv("~/Dropbox/business_insight/df_resp_raw.csv")
resp_df_raw$X1 <- NULL
```